---
title: "LAB 04"
author: "Nana Ama Obeng Nti"
output:
  html_document:
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##QUESTION 1
The number of points scored by each team in a tournament is normally distributed, with
mean = 32 and standard deviation = 7. What is the probability of:

###Question 1a.i
A team scoring 20 points or fewer? 

```{r}
less.20 <- pnorm(20,32,7)
less.20
```

###Question 1a.ii
Plot the density function and shade the area. 

```{r}
curve(dnorm(x, mean = 32, sd = 7), from = 11, to = 53)
pt.x <- c(11,seq(11,20,by=0.01),20)
pt.y <- c(0, dnorm(seq(11,20,by=0.01),mean = 32,sd = 7),0)
polygon(pt.x, pt.y,col = "blue")
```

###Question 1b.i
A team scoring more than 35 points 

```{r}
more.35 <- pnorm(35,32,7,lower.tail = F)
more.35
```

###Question 1b.ii
Plot the density function and shade the area.

```{r}
curve(dnorm(x,32,7),from = 11, to = 53)
pt.x <- c(35,seq(35,53,0.01),53)
pt.y <- c(0,dnorm(seq(35,53,0.01),32,7),0)
polygon(pt.x,pt.y,col = "blue")
```

###Question 1c.i
A team scoring between 20 and 40 points 

```{r}
btwn <- diff(pnorm(c(20,40),32,7))
btwn
```

###Question 1c.ii
Plot the density function and shade the area

```{r}
curve(dnorm(x,32,7),from = 11, to = 53)
pt.x <- c(20,seq(20,40,0.01),40)
pt.y <- c(0, dnorm(seq(20,40,0.01),32,7),0)
polygon(pt.x,pt.y,col = "blue")
```

##QUESTION 2
The number of comments per post on a social media site is exponentially distributed, with the
average post receiving ten comments. What percentage of posts get

###Question 2a.i
Fewer than three comments

```{r}
few.3 <- pexp(3,rate = 1/10)
few.3
```

###Question 2a.ii
Plot the density function and shade the area

```{r}
curve(dexp(x, rate = 1/10),from = 0, to = 100)
pt.x <- c(0,seq(0,3,by=0.01),3)
pt.y <- c(0,dexp(seq(0,3,0.01),rate = 1/10),0)
polygon(pt.x,pt.y,col = "blue")
```

###Question 2b.i
More than 20 comments

```{r}
more.20 <- pexp(20,1/10,lower.tail = F)
more.20
```

###Question 2b.ii
Plot the density function and shade the area

```{r}
curve(dexp(x,1/10),from = 0, to = 100)
pt.x <- c(20,seq(20,100,0.01),100)
pt.y <- c(0,dexp(seq(20,100,0.01),1/10),0)
polygon(pt.x,pt.y,col = "blue")
```

###Question 2c.i
Between five and ten comments

```{r}
btwn <- diff(pexp(c(5,10),1/10))
btwn
```

###Question 2c.ii
Plot the density function and shade the area

```{r}
curve(dexp(x,1/10),from = 0, to = 100)
pt.x <- c(5,seq(5,10,0.01),10)
pt.y <- c(0,dexp(seq(5,10,0.01),1/10),0)
polygon(pt.x,pt.y,col = "blue")
```

##QUESTION 3
Basic raster creation and calculation

###Question 3a.i
Create a first raster (30 rows and 30 columns) and assign values to the raster based on random values from a uniform distribution

```{r, message=FALSE, warning=FALSE}
library(raster)
set.seed(10)
ras.1 <- raster(nrows=30,ncols=30,xmn=0,xmx=30,ymn=0,ymx=30)
ras.1
```

###Question 3a.ii
Plot the raster and the histogram.

```{r}
ras.1[] = runif(ncell(ras.1))
plot(ras.1)
hist(ras.1, main = "RANDOM UNIFORM DISTRIBUTION", col = "blue", density = 15)

```

###Question 3b.i
Create a second raster (30 rows and 30 columns) and assign values to the raster based on random values from a normal distribution

```{r}
set.seed(100)
ras.2 <- raster(nrows=30,ncols=30,xmn=0,xmx=30,ymn=0,ymx=30)
ras.2
```

###Question 3b.ii
Plot the raster and the histogram

```{r}
ras.2[] = rnorm(ncell(ras.2))
plot(ras.2)
hist(ras.2, main = "RANDOM NORMAL DISTRIBUTION", col = "blue", density = 15)
```

###Question 3c.i
Create a third raster (30 rows and 30 columns) and assign values to the raster based on cell-by-cell addition of the first raster and the second raster

```{r}
ras.3 <- ras.1 + ras.2
ras.3
```

###Question 3c.ii
Plot the raster and the histogram

```{r}
plot(ras.3)
hist(ras.3, main = "CELL-BY-CELL ADDITION", col = "blue", density = 15)
```

###Question 3d.i
Calculate the mean value of the third raster. 

```{r}
mean(ras.3)
mn <- cellStats(ras.3,mean)
mn
```

###Question 3d.ii
Reclassify the third raster into a binary image: 1 (cell value > mean value) and 0 (cell value <= mean value)

```{r}
ras.3[ras.3 <= mn] <- 0
ras.3[ras.3 > mn] <- 1
plot(ras.3)
hist(ras.3, main = "RECLASSIFICATION", col = "blue", density = 15)
```

###Question 3d.iii
Save the reclassified image to your working directory as a TIFF image

```{r, message=FALSE, warning=FALSE}
library(rgdal)
writeRaster(ras.3, filename = "ras.3.tif", format = "GTiff", overwrite = T)
read = raster("ras.3.tif")
```


