---
title: "Geog533 Lab 8"
author: "Nana Ama Obeng Nti"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

Complete the following exercises in Chapter 8 (Introduction to Regression Analysis) of the textbook [R] pages 247-250. 

## Question 1

This is Exercise 1 in Chapter 8 of the Textbook [R].

A regression of weekly shopping trip frequency on annual income (data entered in thousands of dollars) is performed on data collected from 24 respondents. The results are summarized below:

* Intercept: 0.46
* Slope: 0.19

|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|                 |                    |                |                 |
| Residual |       1.7       |                    |                |                 |
| Total    |       2.3       |                    |                |                 |


### Question 1(a)
Fill in the blanks in the ANOVA table.

```{r}
n <- 24
SSR <- 1.7
SST <- 2.3
SSM <- SST - SSR

df1 <- 1
df2 <- n - 2
df3 <- n - 1

MMS <- SSM / df1
RMS <-  SSR / df2

F <- MMS / RMS

```


|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|  `r SSM`        |    `r df1`         |  `r MMS`       |    `r F`        |
| Residual |       1.7       |    `r df2`         |  `r RMS`       |                 |
| Total    |       2.3       |  `r df3`           |                |                 |


### Question 1(b)
What is the predicted number of weekly shopping trips for someone making $50,000/year?
```{r}
int <- 0.46
slp <- 0.19
yh <- int + slp * (50000/1000)
yh
```


### Question 1(c)
In words, what is the meaning of the coefficient 0.19?
```{r}
print("Coefficient of 0.19 means that, the predicted weekly shopping trips frequency increases by 0.19 for every increase in annunal income of $1000")
```



### Question 1(d)
Is the regression coefficient significantly different from zero? How do you know?
```{r}
F.crit <- qf(0.95,df1,df2)
F.crit
if(F>F.crit) {
  print("We reject the null hypothesis")
} else {
  print("We cannot the null hypothesis")
}

print("Since the F is greater than F.crit we reject the null hypothesis, thus, the regression coefficient is significantly different from zero")
```



### Question 1(e)
What is the value of the correlation coefficient?
```{r}
R2 <- SSM / SST
R2
r <-  sqrt(R2)
r
```





## Question 2

This is Exercise 6 in Chapter 8 of the Textbook [R].

The following data are collected in an effort to determine whether snowfall is dependent upon elevation: 

| Snowfall (inches) | Elevation (feet) |
|-------------------|------------------|
|       36          |       400        |
|       78          |       800        |
|       11          |       200        |
|       45          |       675        |

Using R, show your work on exercises (a) through (g).

### Question 2(a)
Find the regression coefficients (the intercept and the slope coefficient).

```{r}
elevation <- c(400,800,200,675)
snowfall <- c(36,78,11,45)
Reg <- lm(snowfall~elevation)
summary(Reg)

intercept.slope <- Reg$coefficients
intercept.slope

print("Intercept = -7.63721")
print("Slope = 0.09665")
```


### Question 2(b)
Estimate the standard error of the residuals about the regression line.
```{r}
anova(Reg)
n <- 4
SSR <- 253.23
st.err <- sqrt(SSR / (n-2))
st.err
```


### Question 2(c)
Test the hypothesis that the regression coefficient associated with the independent variables is equal to zero. Also place a 95% confidence interval on the regression coefficient.
```{r}
summary(Reg)
print("Since P-value (0.05663) is greater than 0.05 we accept the null hypothesis that the regression coefficient associated with the independent variables is equal to zero.")

confint(Reg)
```


### Question 2(d)
Find the value of $r^2$.
```{r}
R2 <- summary(Reg)
R2

print("R2 = 0.8899")
```

### Question 2(e)
Make a table of the observed values, predicted values, and the residuals.
```{r}
library(knitr)
df <- data.frame(elevation,snowfall)
df$predict <- Reg$fitted.values
df$residuals <- Reg$residuals
kable(df)
```


### Question 2(f)
Prepare an analysis of variance table portraying the regression results.
```{r}
anv <- anova(Reg)
anv
ssm <- anv["Sum Sq"][[1]][1]
ssr <- anv["Sum Sq"][[1]][2]
sst <- SSM + SSR
df.1 <- anv["Df"][[1]][1]
df.2 <- anv["Df"][[1]][2]
df.3 <- df.1 + df.2
mms <- anv["Mean Sq"][[1]][1]
rms <- anv["Mean Sq"][[1]][2]
F.v <- anv["F value"][[1]][1]

```


|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression| `r ssm`         |`r df.1`            | `r mms`        |`r F.v`          |
| Residual | `r ssr`         | `r df.2`           | `r rms`        |                 |
| Total    |   `r sst`       |`r df.3`            |                |                 |

### Question 2(g)
Graph the data and the regression line.
```{r}
plot(elevation, snowfall)
abline(Reg,col="blue",lwd=3)
```




## Question 3

This is Exercise 10 in Chapter 8 of the Textbook [R].

Use R and the [Milwaukee dataset](http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv) to:

### Problem
perform a regression using sales price as the dependent variable and lot size as the independent variable.

### Solution
```{r}
url = "http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv"
df <- read.csv(url,header = TRUE)
reg <- lm(df$SalePrice~df$LotSize)
summary(reg)
plot(df$LotSize,df$SalePrice)
abline(reg,col="blue",lwd=2)
```



## Question 4

This is Exercise 11 in Chapter 8 of the Textbook [R].

Use R and the Hypothetical [UK Housing Prices dataset](http://spatial.binghamton.edu/geog533/data/UK_Housing.csv) to:

### Question 4(a) 
perform a regression using house price as the dependent variable, and number of bedrooms as the independent variable;

```{r}
url = "http://spatial.binghamton.edu/geog533/data/UK_Housing.csv"
df <- read.csv(url,header = TRUE)
reg <- lm(df$price~df$bedrooms)
summary(reg)
plot(df$bedrooms,df$price)
abline(reg,col="blue",lwd=3)
```


### Question 4(b) 
(b) repeat part (a), using the number of bathrooms as the independent variable, and comment on the results.
```{r}
reg2 <- lm(df$price~df$bathrooms)
summary(reg2)
plot(df$bathrooms,df$price)
abline(reg,col="blue",lwd=3)
```


## Question 5
Download the [hometown.csv]( http://spatial.binghamton.edu/geog533/data/hometown.csv), append your name and hometown coordinates to the csv, then use [leaflet](https://rstudio.github.io/leaflet/) to create a web map (10 pt.)

```{r}
library(leaflet)
df <- read.csv("hometown.csv",header = TRUE)
leaflet(data = df) %>% addTiles() %>% 
  addMarkers(~df$longitude, ~df$latitude, popup = ~as.character(fullname), label = ~as.character(fullname), clusterOptions = markerClusterOptions())
```

