---
title: "Geog533 Lab 7 - Correlation"
author: "Nana Ama Obeng Nti"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

## Question 1
This is Exercise 1 in Chapter 7 of the Textbook [R].

### Problem 1a
Find the correlation coefficient, *r*, for the following sample data on income and education:
```{r}
library(knitr)
obs <- seq(1:5)
income <- c(30,28,52,40,35)
edu <- c(12,13,18,16,17)
df <- data.frame(obs,income,edu)
names(df) <- c("Observation","Income ($*1000)","Education (Years)")
kable(df)
```

Solution 1a
```{r}
r <- cor(df$`Income ($*1000)`, df$`Education (Years)`)
r
```


### Problem 1b
Test the null hypothesis ρ = 0.

Solution 1b
```{r}
r <- cor.test(df$`Income ($*1000)`, df$`Education (Years)`)
r
if(r$p.value>0.05) {
  print("We cannot reject the null hypothesis.")
} else {
  print("We reject the null hypothesis.")
}

print("We cannot reject the null hypothesis that the true correlation is equal to zero because the p.value is greater than 0.05")
```


### Problem 1c
Find Spearman’s rank correlation coefficient for these data.

Solution 1c
```{r}
cor <- cor(df$`Income ($*1000)`, df$`Education (Years)`, method = "spearman")
cor
```


### Problem 1d
Test whether the observed value of rs from part (c) is significantly different from zero.

Solution 1d
```{r}
cor <- cor.test(df$`Income ($*1000)`, df$`Education (Years)`, method = "spearman")
cor
if(cor$p.value>0.05) {
  print("We cannot reject the null hypothesis.")
} else {
  print("We reject the null hypothesis.")
}

print("We cannot reject the null hypothesis that the true correlation is equal to zero because the p.value is greater than 0.05")
```



## Question 2
This is Exercise 3 in Chapter 7 of the Textbook [R].

### Problem
The distribution of the t-statistic for testing the significance of a correlation coefficient has n – 2 degrees of freedom. If the sample size is 36 and α = 0.05, what is the smallest absolute value a correlation coefficient must have to be significant? What if the sample size is 80?

### Solution.a
```{r}
##SAMPLE SIZE IS 36
n <- 36
r.crit <- 2/sqrt(n)
r.crit

print("When the sample size is 36, the smallest absolute value a correlation coefficient must have to be significant is 0.333333")

```
###Solution.b
```{r}
#SAMPLE SIZE IS 80
n <- 80
r.crit <- 2/sqrt(n)
r.crit

print("When the sample size is 80, the smallest absolute value a correlation coefficient must have to be significant is 0.2236068")
```



## Question 3
This is Exercise 4 in Chapter 7 of the Textbook [R].

### Problem
Find the correlation coefficient for the following data:

```{r}
library(knitr)
Obs <- seq(1:4)
X <- c(2,8,9,7)
Y <- c(6,6,10,4)
df <- data.frame(Obs,X,Y)
kable(df)
```

### Solution
```{r}
r.3 <- cor.test(df$X,df$Y,method = "pearson")
r.3
if(r.3$p.value>0.05){
  print("We cannot reject the null hypothesis")
}else{
  print("We reject the null hypothesis")
}

print("We cannot reject the null hypothesis that the true correlation is equal to zero because the p.value is greater than 0.05")
```



## Question 4
This is Exercise 6 in Chapter 7 of the Textbook [R].

### Problem
Find the correlation coefficient between median annual income in the United States and the number of horse races won by the leading jockey, for the period 1984–1995. Test the hypothesis that the true correlation coefficient is equal to zero. Interpret your results.

```{r}
year <- 1984:1994
income <- c(35165,35778,37027,37256,37512,37997,37343,36054,35593,35241,35486)
races <- c(399,469,429,450,474,598,364,430,433,410,317)
df <- data.frame(year,income,races)
names(df) <- c("Year","Median income","Number of races won by leading jockey")
kable(df)
```


### Solution
```{r}
r.4 <- cor.test(df$`Median income`,df$`Number of races won by leading jockey`)
r.4
if(r.4$p.value>0.05){
  print("We cannot reject the null hypothesis")
}else{
  print("We reject the null hypothesis")
}

print("We cannot reject the null hypothesis that the true correlation is equal to zero because the p.value is greater than 0.05")

##INTERPRET YOUR RESULTS
n <- 11
r.crit <- 2/sqrt(n)
r.crit

print("Eventhough the null hypothesis is accepted, the correlation coefficient of 0.558 is not significant. This is because, a sample size of 11 must have a correlation coefficient of at least 0.603 to be considered significant.")
```



## Question 5
This is Exercise 7 in Chapter 7 of the Textbook [R].

### Problem
For the following ranked data, find Spearman’s r, and then test the null hypothesis (using a Type I error probability of 0.10) that the true correlation is equal to zero.

```{r}
library(knitr)
obs <- 1:6
x <- c(1,2,5,6,11,12)
y <- c(8,4,12,3,10,7)
df <- data.frame(obs,x,y)
names(df) <- c("Observation","Rank of x","Rank of y")
kable(df)
```


### Solution
```{r}
r.5 <- cor.test(df$`Rank of x`,df$`Rank of y`,method = "spearman",conf.level = 0.90)
r.5
if(r.5$p.value>0.1){
  print("We cannot reject the null hypothesis")
}else{
  print("We reject the null hypothesis")
}

print("We cannot reject the null hypothesis that the true correlation is equal to zero because the p.value is greater than 0.05")
```



## Question 6
This is Exercise 8 in Chapter 7 of the Textbook [R].

### Problem
Find Pearson’s r for the following data, and then test the null hypothesis that the correlation coefficient is equal to zero. Use a Type I error probability of 0.05.

```{r}
library(knitr)
obs <- 1:6
x <- c(3.2,2.4,1.6,8.3,7.2,5.1)
y <- c(6.2,7.3,8.1,2.6,6.3,4.3)
df <- data.frame(obs,x,y)
names(df) <- c("Observation","x","y")
kable(df)
```


### Solution
```{r}
r.6 <- cor.test(df$x,df$y,method = "pearson")
r.6
if(r.6$p.value>0.05){
  print("We cannot reject the null hypothesis")
}else{
  print("We reject the null hypothesis")
}

print("We cannot reject the null hypothesis that the true correlation is equal to zero because the p.value is greater than 0.05")
```


## Question 7
This is Exercise 9 in Chapter 7 of the Textbook [R].

### Problem
Using R and the Milwaukee dataset, find the correlation between number of bedrooms and lot size.

### Solution
```{r}
df <- read.csv("Milwaukee_Sales.csv",header = T)
df
r.7 <- cor.test(df$Bedrms,df$LotSize)
r.7
if(r.7$p.value>0.05){
  print("We cannot reject the null hypothesis")
}else{
  print("We reject the null hypothesis")
}

print("We cannot reject the null hypothesis that the true correlation is equal to zero because the p.value is greater than 0.05")
```


## Question 8
This is Exercise 10 in Chapter 7 of the Textbook [R].

### Problem
Using R and the Hypothetical UK Housing Prices dataset, find the correlation between floor area and number of bedrooms.

### Solution
```{r}
df <- read.csv("UK_Housing.csv",header = T)
df
r.8 <- cor.test(df$floorarea,df$bedrooms)
r.8
if(r.8$p.value>0.05){
  print("We cannot reject the null hypothesis")
}else{
  print("We reject the null hypothesis")
}

print("We reject the null hypothesis that the true correlation is equal to zero because the p.value is less than 0.05")
```


## Question 9
Use the **cars** data frame in the **datasets** package to perform the following tasks: 

### Problem 9a
Plot a scatterplot for the data frame (x: speed, y: dist)

Solution 9a
```{r}
library(datasets)
df <- cars
plot(df,xlab="Speed",ylab="dist",col="blue")
```


### Problem 9b
How many rows in the data frame?

Solution 9b
```{r}
n <- nrow(df)
n
```


### Problem 9c
Calculate Pearson’s correlation coefficient using the equation below:
$$r=\frac{\sum_{n}^{i=1}(x_i-\bar{x})(y_i-\bar{y})}{(n-1)s_{x}s_{y}}$$

Solution 9c
```{r}
library(datasets)
df <- cars

speed <- df$speed
spd.mn<- mean(speed)
spd.mn
dist <- df$dist
dis.mn <- mean(dist)
dis.mn

sd1 <- sd(speed)
sd1
sd2 <- sd(dist)
sd <- sd1 * sd2
sd
n <- 50
n1 <- n - 1
n1

num <- sum((speed - spd.mn)*(dist - dis.mn))
num
deno <-  n1 * sd
deno

r <- num / deno
r

t <- r * sqrt (n - 2) / sqrt(1-r^2)
t

t.crit <- qt(0.975, df = n-2)
t.crit

print("We reject the null hypothesis that the true correlation is equal to zero, since the observed value of t (9.46) is greater than the critical value of 2.01")
```


### Problem 9d
Use the cor.test() function to find Pearson’s correlation coefficient and compare it to the one from part (c)

Solution 9d
```{r}
r <- cor(df$speed, df$dist, method = "pearson")
r
result <- cor.test(df$speed,df$dist, method = "pearson")
result
if(result$p.value>0.05){
  print("We cannot reject the null hypothesis")
}else{
  print("We reject the null hypothesis")
}

print("We reject the null hypothesis that the true correlation is equal to zero because the p.value is less than 0.05")

print("9d and 9c produce the same correlation coefficient output")
```


### Problem 9e
Calculate Spearman’s rank correlation coefficient using the equation below:
$$r_{S} = 1 - \frac{6\sum_{i=1}^{n}d_{i}^{2}}{n^3-n}$$

Solution 9e
```{r}
library(datasets)
df <- cars
df
n <- nrow(df)

df$rank.sp <- rank(df$speed)
df$rank.ds <- rank(df$dist)

df$d <- df$rank.sp - df$rank.ds
df$d2 <- df$d^2

sum.d2 <- sum(df$d2)
sum.d2

num <- 6 * sum.d2
num

deno <- n^3 - n
deno

x <- num / deno
x

rs <- 1 - x
rs
```



### Problem 9f
Use the cor.test() function to find Spearman’s rank correlation coefficient and compare it to the one from part (e)

Solution 9f
```{r}
rs <- cor(df$speed, df$dist, method = "spearman")
rs
result1 <- cor.test(df$speed,df$dist, method = "spearman")
result1
if(result1$p.value>0.05){
  print("We cannot reject the null hypothesis")
}else{
  print("We reject the null hypothesis")
}

print("We reject the null hypothesis that the true correlation is equal to zero because the p.value is less than 0.05")

print("9e and 9f produce the same correlation coefficient output")
```

