---
title: "Geog533 Lab 10"
author: "Nana Ama Obeng Nti"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

## Question 1
Use R package UScensus2010county to complete the following tasks:  (20 pt.)

### Question 1(a)
Plot a map of New York counties using the plot function.
```{r echo=TRUE, message=FALSE, warning=FALSE}
if (!require(UScensus2010)) install.packages("UScensus2010")
library(UScensus2010)
if(!require(UScensus2010county)) install.county("osx")
if(!require(UScensus2000tract)) install.tract("osx")
library(UScensus2010county)
library(UScensus2010tract)


data("new_york.county10")
county <- new_york.county10
plot(county)
```

### Question 1(b)	
Plot a map of New York counties using the qtm function.
```{r echo=TRUE, message=FALSE, warning=FALSE}
if(!require(tmap)) install.packages("tmap")
library(tmap)
qtm(shp = county,fill = "NAME10",fill.title="County",style="col_blind")
```


### Question 1(c)	
How many counties in New York State?
```{r message=FALSE, warning=FALSE}
df <- county@data
print(nrow((df)))
```

### Question 1(d)	
What’s the 3-digit fips code of Broome County?
```{r}
Broome <- county.df[df$NAME10=="Broome",]
print(Broome$fips)
```

### Question 1(e)	
Compute descriptive statistics of the population column (P0010001), including total, minimum, maximum, mean, median, and skewness. 
```{r echo=TRUE, message=FALSE, warning=FALSE}
if(!require(fBasics)) install.packages("fBasics")
library(fBasics)
library(fBasics)
pop <- df$P0010001
des.stats <- basicStats(pop)
des.stats["Sum",]
des.stats["Minimum",]
des.stats["Maximum",]
des.stats["Mean",]
des.stats["Median",]
des.stats["Skewness",]

```

### Question 1(f)	
Create a histogram and a boxplot of the population.
```{r}
hist(pop)
boxplot(pop)
```


## Question 2
Use R package UScensus2010tract to complete the following tasks:    (20 pt.)
```{r echo=TRUE, message=FALSE, warning=FALSE}
library(UScensus2010tract)
```

### Question 2(a)	
Plot a map of New York census tracts using the plot function.
```{r}
data("new_york.tract10")
tract <- new_york.tract10
plot(tract)
```

### Question 2(b)
Compute the total population based on census tracts.
```{r}
df2 <- tract@data
tract.pop <- df2$P0010001
tract.stats <- basicStats(tract.pop)
tract.stats["Sum",]
```

### Question 2(c)
Select all census tracts in Broome County and plot the map. 
```{r}
tracts.Broome <- tract[df2$county=="007",]
plot(tracts.Broome)
```

### Question 2(d)
What’s the total population of Broome County?
```{r}
pop.brm <- tracts.Broome$P0010001
t.pop <- basicStats(pop.brm)
t.pop["Sum",]
```

### Question 2(e)
Create a histogram and a boxplot of population based on census tracts of Broome County.
```{r}
hist(pop.brm)
boxplot(pop.brm)
```

### Question 2(f)
Select the first five columns of the shapefile of Broome County census tract; add a new population ratio column (= census tract population / county population); save the new shapefile to the hard drive. 
```{r echo=TRUE, message=FALSE, warning=FALSE}
Broome <- tracts.Broome[,1:5]
df3 <- Broome@data
df3$pop.ratio <- df3$P0010001/{sum(pop.brm)}
Broome@data <- df3
if(!require(rgdal)) install.packages("rgdal")
library(rgdal)
library(rgdal)
writeOGR(Broome,dsn = "output",layer = "Broome",driver = "ESRI Shapefile",overwrite_layer = TRUE)
print(df3)
```


## Question 3

Use R packages raster and ncdf4 to complete the following tasks:     (20 pt.)
```{r echo=TRUE, message=FALSE, warning=FALSE}
if(!require(ncdf4)) install.packages("ncdf4")
library(raster)
library(ncdf4)
```

### Question 3(a)		
Load the multi-band raster dataset “NDVI.nc” into R.
```{r}
NDVI <- brick("NDVI.nc")
NDVI
```

### Question 3(b)		
Get the basic information about the dataset, including the number of rows, columns, cells, and bands; spatial resolution, extent, bounding box, and projection.
```{r}
nrow(NDVI)
ncol(NDVI)
ncell(NDVI)
nlayers(NDVI)
res(NDVI)
extent(NDVI)
bbox(NDVI)
projection(NDVI)
```

### Question 3(c)	
Aggregate all bands to generate a mean NDVI raster and a maximum NDVI raster; save the two new raster datasets to the hard drive. 
```{r message=FALSE, warning=FALSE}
NDVI.mean <- mean(NDVI)
plot(NDVI.mean)

NDVI.max <- max(NDVI)
plot(NDVI.max)

writeRaster(NDVI.mean,filename = "output/NDVI.mean.tif",overwrite=TRUE)
writeRaster(NDVI.max,filename = "output/NDVI.max.tif",overwrite=TRUE)
```

### Question 3(d)			
Plot the maps of monthly NDVI of the year 2001
```{r message=FALSE, warning=FALSE}
plot(NDVI,13:24)
```

### Question 3(e)	
Create histograms of monthly NDVI of the year 2001.
```{r message=FALSE, warning=FALSE}
hist(NDVI,13:24)
```

### Question 3(f)			
Plot the NDVI map of July 2000; click any location with data on the map and retrieve the NDVI time series for all years; plot the NDVI time series of the selected location. 
```{r message=FALSE, warning=FALSE}
plot(NDVI,1)
values <- click(NDVI, n=1, xy=TRUE)	
values <- click(NDVI, n=1, xy=FALSE)
values <- NDVI[50,50]
plot(as.vector(values), type="b")
```

