---
title: "LAB 02"
author: "Nana Ama Obeng Nti"
date: "September 23, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(MASS)
cars <- Cars93
```


## QUESTION 1a

Manufacturer and Model for cheapest car (Price) based of car types

```{r}
Price.min <- tapply(cars$Price, cars$Type, min)
Price.min

cheap.price <- cars[c(74, 20, 47, 31, 46, 16), c(1:3, 5)]
cheap.price
```

## QUESTION 1b

Manufacturer  and Model for car with the greatest fuel efficiency (MPG.highway) for each car type

```{r}
MPG.max <- tapply(cars$MPG.highway, cars$Type, max)
MPG.max

max <- cars[c(12,7,71,77,20,6,69,39,40,56), c(1:3,8)]
max
```

## QUESTION 2a

Mean Horsepower for each car type

```{r}
Horsepower.mean <- tapply(cars$Horsepower, cars$Type, mean)
Horsepower.mean
```

## QUESTION 2b

Differnce between each cars Horsepower and mean for its type. ii) Skewness and Kurtosis for each car type based on the difference.

```{r, include=FALSE}
library(moments)
```


```{r}
cars$Horsepower.mean <- Horsepower.mean[cars$Type]
cars$Horsepower.diff <- cars$Horsepower - cars$Horsepower.mean
skewness(cars$Horsepower.diff)
kurtosis(cars$Horsepower.diff)
```


## QUESTION 3

Dataframes for USA and non-USA cars

```{r}
USA.df <- cars[cars$Origin == "USA", ]
USA.df

NUSA.df <- cars[cars$Origin == "non-USA", ]
NUSA.df
```

## QUESTION 4

Saving USA cars to file with write.csv()

```{r}
write.csv(USA.df, file = "USA.csv", row.names = FALSE)

read.csv("USA.csv")
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
